### 2014-04-26 stat246/hw3.tex
s3

We know that $\kl(X\|U)\geq0$, with equality when $X=U$, so the upper bound for $H(X)$ is $\log K$. A lower bound for the entropy is $0$, since the log term will always be non-positive and the probability term will always be non-negative. This is a tight bound, since if we let $P(X=1)\to1$ and the other probabilities tend towards zero, since $\lim_{x\to0}\log x^x=0$ (for the $k\neq1$ terms) and $\lim_{x\to1}\log x^x=0$ (for the $k=1$ term).
\ssn{b,c}
\begin{align*}
    H(X,Y)&=-\sum_{i=1}^K\sum_{j=1}^KP(i,j)\log P(i,j)\\
          &=-\sum_{i=1}^K\sum_{j=1}^KP(i|j)P(j)\log \left(P(i|j)P(j)\right)\\
          &=-\sum_{j=1}^KP(j)\sum_{i=1}^KP(i|j)\left(\log P(i|j) + \log P(j)\right)\\
          &=-\sum_{j=1}^KP(j)\left(\sum_{i=1}^KP(i|j)\log P(i|j)+\log P(j)\sum_{i=1}^KP(i|j)\right)\\
          &=H(X|Y)-\sum_{j=1}^KP(j)\log P(j)\\
          &=H(X|Y)+H(Y)\\
\end{align*}
Huh. Looks like I'm doing this backwards then. Writing $H(X|Y)$ as $\sum_{i=1}^K\sum_{j=1}^KP(i,j)\log \P(i|j)$, we have 
\begin{align*}
    H(X|Y)-H(X)&=\sum_{i=1}^KP(i)\log P(i)-\sum_{i=1}^K\sum_{j=1}^KP(i,j)\log P(i|j)\\
               &=\sum_{i=1}^K \left(P(i)\log P(i)-\sum_{j=1}^KP(i,j)\log P(i|j)\right)\\
               &=\sum_{i=1}^K\sum_{j=1}^KP(i,j)\log\frac{P(i)}{P(i|j)}\\
               &\leq\log\left(\sum_{i=1}^K\sum_{j=1}^K\frac{P(i,j)P(i)}{P(i|j)}\ \text{by Jensen's ineq}\\
               &=\log\left(\sum_{i=1}^K\sum_{j=1}^KP(i)P(j)\right)\\
               &=0
\end{align*}
so we have $H(X|Y)\leq H(X)$, with equality only when $\frac{P(i)}{P(i|j)}$ are equal for every $i,j$. This only happens when $P(i|j)=P(i)$ for all $i,j$ ($X$ and $Y$ independent), since otherwise we'd have some $i$ for which there exists $j_1$ and $j_2$ such that $P(i|j_1)>P(i)$ and $P(i|j_2)<P(i)$. Then, it follows that $H(X,Y)=H(X|Y)+H(Y)\leq H(X)+H(Y)$ with equality iff $X$ and $Y$ are independent.

>>>>

We know that $\kl(X\|U)\geq0$, with equality when $X=U$, so the upper bound for $H(X)$ is $\log K$. A lower bound for the entropy is $0$, since the log term will always be non-positive and the probability term will always be non-negative. This is a tight bound, since if we let $P(X=1)\to1$ and the other probabilities tend towards zero, since $\lim_{x\to0}\log x^x=0$ (for the $k\neq1$ terms) and $\lim_{x\to1}\log x^x=0$ (for the $k=1$ term).
\ssn{b,c}
\begin{align*}
    H(X,Y)&=-\sum_{i=1}^K\sum_{j=1}^KP(i,j)\log P(i,j)\\
          &=-\sum_{i=1}^K\sum_{j=1}^KP(i|j)P(j)\log \left(P(i|j)P(j)\right)\\
          &=-\sum_{j=1}^KP(j)\sum_{i=1}^KP(i|j)\left(\log P(i|j) + \log P(j)\right)\\
          &=-\sum_{j=1}^KP(j)\left(\sum_{i=1}^KP(i|j)\log P(i|j)+\log P(j)\sum_{i=1}^KP(i|j)\right)\\
          &=H(X|Y)-\sum_{j=1}^KP(j)\log P(j)\\
          &=H(X|Y)+H(Y)\\
\end{align*}
Huh. Looks like I'm doing this backwards then. Writing $H(X|Y)$ as $\sum_{i=1}^K\sum_{j=1}^KP(i,j)\log \P(i|j)$, we have 
\begin{align*}
    H(X|Y)-H(X)&=\sum_{i=1}^KP(i)\log P(i)-\sum_{i=1}^K\sum_{j=1}^KP(i,j)\log P(i|j)\\
               &=\sum_{i=1}^K \left(P(i)\log P(i)-\sum_{j=1}^KP(i,j)\log P(i|j)\right)\\
               &=\sum_{i=1}^K\sum_{j=1}^KP(i,j)\log\frac{P(i)}{P(i|j)}\\
               &\leq\log\left(\sum_{i=1}^K\sum_{j=1}^K\frac{P(i,j)P(i)}{P(i|j)}\right)\ \text{by Jensen's ineq}\\
               &=\log\left(\sum_{i=1}^K\sum_{j=1}^KP(i)P(j)\right)\\
               &=0
\end{align*}
so we have $H(X|Y)\leq H(X)$, with equality only when $\frac{P(i)}{P(i|j)}$ are equal for every $i,j$. This only happens when $P(i|j)=P(i)$ for all $i,j$ ($X$ and $Y$ independent), since otherwise we'd have some $i$ for which there exists $j_1$ and $j_2$ such that $P(i|j_1)>P(i)$ and $P(i|j_2)<P(i)$. Then, it follows that $H(X,Y)=H(X|Y)+H(Y)\leq H(X)+H(Y)$ with equality iff $X$ and $Y$ are independent.
